\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-21 notes/2loopNLSM.tex
Resized YM numerator span diagrams.
fixed forward refs in catalog app

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}\label{cubicCKLag}
\mathcal{L} = P_V^{ab}P_W^{\mu\nu}\mathcal{O}^a_\mu \Box \mathcal{O}^b_\nu +  V^{abc} W^{\mu \nu \rho}\mathcal{O}^a_\mu \mathcal{O}^b_\nu \mathcal{O}^c_\rho
\end{equation}
where $V^{abc}$ and $W^{\mu \nu \rho}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_\mu$ indexed by quantum numbers $a$ and
$\mu$. The Feynman rules for this theory are simply:
\be\label{CKFrules}
\begin{split}
\propFrule{$a,\mu$}{$b,\nu$}{$k \rightarrow $}{} &= \frac{i}{k^2} P_V^{ab}P_W^{\mu\nu}
\\
\cubicFrule{$a,\mu$}{$b,\nu$}{$c,\rho$}{}{}{} &= i V^{abc}W^{\mu \nu \rho}
\end{split}
\ee
We have introduced $P_V^{ab}$ and $P_W^{\mu\nu}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction, as is the case for $DF^2$ theory \cite{Johansson:2017srf}. Though generally these quadratic dressings are simply delta functions, $P_V^{ab} = \delta^{ab}$, or flat space metrics, $P_W^{\mu\nu}=\eta^{\mu\nu}$. As written, the theory will be color dual if the off-shell cubic vertex is antisymmetric and satisfies the Jacobi identity,
specifically,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0
\end{align}
where using the Feynman rules of \cref{CKFrules}, the bracketed expression is the half-ladder numerator of the following diagram,
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv N\left[\halfLadder\right]\,,
\end{equation}
and likewise for the $W^{\mu \nu \rho}$ dressing. Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional operators \cite{Carrasco:2022lbm,Carrasco:2022sck}, sometimes including additional propagating fields not originally considered in the spectrum. See ref.~\cite{Ben-Shahar:2022ixa} for a recent
example. Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic two-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

Every numerator is required to respect all of the symmetries (or
automorphisms) of its associated graph.  Graph symmetries are an
avatar of Bose symmetry.  For a pure ansatz construction it is very
reasonable to impose graph symmetries, but color-kinematics duality at
tree-level is often achieved at the expense of manifest Bose symmetry
\cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf,
  Edison:2020ehu}.  In an ansatz construction graph symmetries are
imposed for \emph{every} graph, not just the basis graphs.
For example, the symmetries of the crossed box are spanned by,
\begin{equation}
\scalebox{0.8}{\CrossedBoxGraphSym{1}{2}{3}{4}}\!\! \sim\!\! \scalebox{0.8}{\CrossedBoxGraphSym{2}{1}{3}{4}}
\text{ and }
\scalebox{0.8}{\CrossedBoxGraphSym{1}{2}{3}{4}} \!\!\sim\!\! \scalebox{0.8}{\CrossedBoxGraphSym{2}{1}{4}{3}}
\end{equation}
% \ace{is there a reason we commented out this diagram?} \jm{You can try
 % either one.  I liked the one I left in better.}  For example, the
%symmetries of the crossed box
%\begin{equation}
%\CrossedBoxGraphSym{1}{2}{3}{4}
%\end{equation}
which are just functional relabelings of external legs,
$\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.  In general,
internal edge labels must be tracked as well.  Furthermore, care must
be taken with the signs in a symmetry relation.  To compensate for the
antisymmetric color factors, $f^{abc}$, graph vertices are all totally
antisymmetric and hence have signs built into their orientations.
When used in conjuction with the Jacobi relations, \eg
\cref{eq:TwoLoopJacobiOne}, the symmetry constraints of non-basis
graphs still impose conditions on the basis ansatz.  In fact, since
the symmetry properites of a graph are directly linked with
transformations of its color dressing, the resulting constraints can
be thought of as a subclass of the \emph{boundary} Jacobi relations.

@@

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge.
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, J-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-Abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the J-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply:
\be
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}
\ee
As one can see, the kinetic term mixes the two gauge fields and has the standard
Feynman rule of $i/k^2$.  We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the abelian polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.

@@

Below we review the four manifestly color-dual theories that appear in
this paper, in particular SDYM, two formulations of NLSM, and ZM.  All
of these theories are closely tied together by the semi-abelian YM
theory in \cref{eq:semiYM} or in other words the cubic sector of YM.
The manifestly color-dual equations of motion for SDYM and NLSM are
derived by picking a gauge and then placing a constraint not on the YM
equation of motion or Bianchi identity but rather directly on the
field strength itself.  Other theories can be obtained from SDYM and
NLSM by judicious modification of the propagator.  In all of these
theories, the kinematic algebra is that of (volume preserving)
diffeomorphisms or some close variant thereof.  In almost all of these
theories, Bose symmetry is broken or obscured at the Lagrangian level.
Because of the broken Bose symmetry (and the inherited kinetic mixing
of semi-abelian YM), these theories cannot produce color-dual
numerators at two-loop or higher.

\ace{There are some cyclic references in the catalog that are gonna be
  a bit annoying to clean up: ZM refers to J, YZ refers to ZM and
  SDYM, J refers to YZ.  I have currently switched the order to: SDYM,
  ZM, YZ, J, but we should probably discuss.}

\subsection{Self-Dual Yang-Mills} The first example in the literature
of color-dual Feynman rules is that of Yang-Mills theory in the
self-dual sector (SDYM).  We summarize the appoach of
Ref.~\cite{Monteiro2011pc} for extracting the kinematic algebra.
% \begin{equation}
% \mathcal{L}^{\text{YM}} = -\frac{1}{4}F_{\mu\nu}F^{\mu\nu}
% \end{equation}
To obtain the Lagrangian for self-dual Yang-Mills from standard Yang-Mills, we apply the
self-dual condition that equates the non-abelian $SU(N)$ field
strength,
$F_{\mu\nu} = \partial_\mu A_\nu -\partial_\nu A_\mu + g[A_\mu
,A_\nu]$, to its dual two-form:
\begin{equation}
  \label{SDcon} F_{\mu\nu} = \frac{1}{2}
  \varepsilon_{\mu\nu\rho\sigma}F^{\rho\sigma} \,.
\end{equation}
Solutions to the self-dual condition automatically satisfy the
Yang-Mills equation of motion due to the Bianchi identity:
\begin{equation}
D^\mu F^{\mu\nu} = \frac{1}{2} \varepsilon_{\mu\nu\rho\sigma}D^\mu F^{\rho\sigma} = 0 \,.
\end{equation}
Solutions to \cref{SDcon} correspond to instanton configurations that
are probed by the gauge field path integral in the full
non-perturbative description of Yang-Mills. In this paper, we will
focus on the perturbative aspects of this constraint. We begin by
transforming from Cartesian coordinates to light cone coordinates,
$(t,x,y,z) \rightarrow (u,v,w,\bar{w})$, via:
\begin{equation}
u = t+x,\quad v=t-x, \quad w = y+iz
\end{equation}
which yields the invariant line element:
\begin{equation}
ds^2 = 2(dudv - dwd\bar{w}) = dt^2 - d\vec{x}\cdot d\vec{x} \,.
\end{equation}
Applying this coordinate transformation to the self-dual condition of
\cref{SDcon}, produces three simple constraint equations:
\begin{align} \label{SDLightCone1}
F_{uw} &=0\,,
\\
 \label{SDLightCone2}
 F_{uv} &= F_{w\bar{w}}\,,
\\
 \label{SDLightCone3}
 F_{v\bar{w}} &= 0\,.
\end{align}
Thus, in light-cone gauge, $A_u=0$, we realize the self-dual
conditions in terms of the gauge field as
\begin{align}
  A_u=0 \text{ and } (\ref{SDLightCone1}) \quad&\Rightarrow \quad A_w=0\\
  A_w=0\,,\,A_u=0 \text{ and } (\ref{SDLightCone2}) \quad&\Rightarrow \quad \partial_u A_v = \partial_w A_{\bar{w}} \label{eq:sd-divs}\,.
\end{align}
This last constraint can be manifested by writing the gauge field in
terms of a single degree of freedom which we call $\Psi$:
\begin{equation}
  \label{SDDef}
A_v = \frac{1}{2} \partial_w \Psi, \qquad A_{\bar{w}} = \frac{1}{2} \partial_u \Psi \,.
\end{equation}
Using this definition on the first self-dual constraint equation
yields the following equation of motion for the dynamical scalar
field:
\begin{equation}
  \label{SDEOM}
(\ref{SDLightCone1}) \text{ and } (\ref{SDDef}) \quad  \Rightarrow\quad \Box \Psi + ig [\partial_u \Psi, \partial_w \Psi ]=0 \,.
\end{equation}
Adding back in the anti-holomorphic field $\bar{\Psi}$ as a Lagrange
multiplier, we obtain the following interaction Lagrangian for
self-dual Yang-Mills theory:
\begin{equation}
  \mathcal{L}^{\text{SDYM}} = (\partial \bar{\Psi})(\partial \Psi) -i g \bar{\Psi} [\partial_u \Psi, \partial_w \Psi ] \,.
\end{equation}
The color-ordered ordered cubic Feynman rule for this theory can be
immediately read off as:
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =X(k_1,k_2) \equiv k_{u,1}k_{w,2} - k_{w,1}k_{u,2} \,.
\end{equation}
In this form, the Lagrangian and Feynman rules are manifestly
color-dual \textit{off-shell}. We can see the effect \cref{SDcon} had
on the Feynman rules of the full Yang-Mills Lagrangian: it simply
decoupled the anti-MHV three-point vertex from the theory, leaving
only the MHV three-point vertex. Before proceeding, it is important to
note that the mass-dimension of the theory appears to differ from that
of Yang-Mills -- we comment on this in detail in \cref{Conventions}.

\subsection{Zakharov-Mikhailov theory}
\label{sec:ZMTheory}
While J-theory manifests color-kinematics duality, it only does so at
tree level because J-theory breaks manifest Bose symmetry.  Manifest
color-kinematics duality can be achieved at all loop orders by
restricting J-theory to two spacetime dimensions where the chiral
current can be dualized to a scalar
$J^\mu = \varepsilon^{\mu\nu} \partial_\nu \varphi$, which trivially
enforces Lorenz gauge.  After plugging $\varphi$ into
\cref{eq:JTheoryEOM} and rearranging, another copy of $\varphi$
(rather than some new field) can be integrated in to obtain the
Zakharov-Mikhailov (ZM) Lagrangian
\begin{equation}
\label{eq:ZMLagrangian}
\mathcal{L}^{\text{ZM}} = \frac{1}{2}(\partial \varphi)^2 + g f^{abc} \varphi^a \varepsilon^{\mu\nu}(\partial_\mu \varphi^b)( \partial_\nu \varphi^c) =  \text{tr}\left[\frac{1}{2}(\partial \varphi)^2-i g \varphi [\partial_t \varphi, \partial_z \varphi ]\right] .
\end{equation}
This procedure makes it clear that ZM and NLSM encode the same
classical physics, that is, their equations of motion are dual to each
other.  ZM theory can also be tenuously motivated from SDYM theory.
The power counting of the interaction term in the SDYM equation of
motion \cref{SDEOM} makes it clear that the theory is critical in 2D.
This suggests that the two lightcone coordinates $u$ and $w$ should be
identified as the coordinates of the 2D space.  The 4D d'Alembertian
in lightcone coordinates,
$\Box = 4 (\partial_u \partial_v - \partial_w \partial_{\bar{w}})$,
cannot be interpreted as the 2D equivalent, so the kinetic term of
SDYM must be replaced with the correct 2D version.  This argument
should be understood purely as an alternative route to ZM theory.
While NLSM and ZM are the same theory classically when restricted to
2D, SDYM and ZM are not because, for example, the tree amplitudes of
SDYM vanish but those of ZM do not (cite argumentative papers).

>>>>

\subsection{Color-dual Lagrangians}\label{offShellCK}
While the BCJ relations are an on-shell statement about the color-dual
nature of scattering amplitudes, one might aspire to hard code
color-kinematics duality by realizing an off-shell Lagrangian
description of the theory. The double-copy construction of
\cref{DCconstruction} would suggest a manifestly cubic Lagrangian of
the form
\begin{equation}\label{cubicCKLag}
\mathcal{L} = P_V^{ab}P_W^{\mu\nu}\mathcal{O}^a_\mu \Box \mathcal{O}^b_\nu +  V^{abc} W^{\mu \nu \rho}\mathcal{O}^a_\mu \mathcal{O}^b_\nu \mathcal{O}^c_\rho
\end{equation}
where $V^{abc}$ and $W^{\mu \nu \rho}$ are cubic Feynman rules mixing field
operators $\mathcal{O}^{a}_\mu$ indexed by quantum numbers $a$ and
$\mu$. The Feynman rules for this theory are simply:
\be\label{CKFrules}
\begin{split}
\propFrule{$a,\mu$}{$b,\nu$}{$k \rightarrow $}{} &= \frac{i}{k^2} P_V^{ab}P_W^{\mu\nu}
\\
\cubicFrule{$a,\mu$}{$b,\nu$}{$c,\rho$}{}{}{} &= i V^{abc}W^{\mu \nu \rho}
\end{split}
\ee
We have introduced $P_V^{ab}$ and $P_W^{\mu\nu}$ projection operators
to encode non-local kinetic structure that could in principle
participate in the construction, as is the case for $DF^2$ theory \cite{Johansson:2017srf}. Though generally these quadratic dressings are simply delta functions, $P_V^{ab} = \delta^{ab}$, or flat space metrics, $P_W^{\mu\nu}=\eta^{\mu\nu}$. As written, the theory will be color dual if the off-shell cubic vertex is antisymmetric and satisfies the Jacobi identity.
Specifically,
\begin{align}
\textit{anti-symmetry}:& \quad{}_{a_1}\langle V^{a_2} \rangle_{a_3}+\text{cyc}(23)=0
\\
\textit{Jacobi identity}:& \quad {}_{a_1}\langle V^{a_2} V^{a_3}\rangle_{a_4}+\text{cyc}(234)=0\label{jacID}
\end{align}
where using the Feynman rules of \cref{CKFrules}, the bracketed expression is the half-ladder numerator of the following diagram,
\begin{equation}
{}_{a_1}\langle V^{a_2}V^{a_3}\cdots V^{a_{n-1}} \rangle_{a_n} \equiv N\left[\halfLadder\right]\,,
\end{equation}
and likewise for the $W^{\mu \nu \rho}$ dressing. Off-shell descriptions of
color-dual theories are rare, and only a few examples are available in
the literature \cite{Monteiro2011pc, Cheung:2016prv, Cheung:2021zvb,
  Cheung:2020djz, Ben-Shahar:2022ixa, Ben-Shahar:2021zww,
  Ben-Shahar:2021doh}.
Typically when one goes about constructing color-dual
theories directly from a set of cubic interactions, higher point
Jacobi and antisymmetry constraints require introducing additional operators \cite{Carrasco:2022lbm,Carrasco:2022sck}, sometimes including additional propagating fields not originally considered in the spectrum. See ref.~\cite{Ben-Shahar:2022ixa} for a recent
example. Indeed, as we will show in \cref{sec:cubic}, Yang-Mills
requires introducing a cubic two-form interaction in order to manifest
the kinematic algebra just at four-point. Absent information about the
kinematic algebra, one can make progress in the construction of
loop-level color-dual amplitudes by employing a color-dual bootstrap.

@@

Every numerator is required to respect all of the symmetries (or
automorphisms) of its associated graph.  Graph symmetries are an
avatar of Bose symmetry.  For a pure ansatz construction it is very
reasonable to impose graph symmetries, but color-kinematics duality at
tree-level is often achieved at the expense of manifest Bose symmetry
\cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf,
  Edison:2020ehu}.  In an ansatz construction graph symmetries are
imposed for \emph{every} graph, not just the basis graphs.
For example, the symmetries of the crossed box are spanned by,
\nhp{Alex, fix the scaling to get the boxes to fit}
\be
\CrossedBoxGraphSym{1}{2}{3}{4}\!\! \sim\!\! \CrossedBoxGraphSym{2}{1}{3}{4} 
\text{ and }
\CrossedBoxGraphSym{1}{2}{3}{4} \!\!\sim\!\! \CrossedBoxGraphSym{2}{1}{4}{3}
\ee
%\ace{is there a reason we commented out this diagram?} \jm{You can try
 % either one.  I liked the one I left in better.}  For example, the
%symmetries of the crossed box
%\begin{equation}
%\CrossedBoxGraphSym{1}{2}{3}{4}
%\end{equation}
which are just functional relabelings of external legs, $\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.
In general, internal edge labels must be tracked as well.
Furthermore, care must be taken with the signs in a symmetry relation.
To compensate for the antisymmetric color factors, $f^{abc}$, graph
vertices are all totally antisymmetric and hence have signs built into
their orientations.  Once imposing symmetry relations, the crossed box can be expressed in terms of the double box
and penta-triangle, thus imposing non-trivial constraints on the
ansatz. 
\nhp{Someone add in comment here about how symmetry relations imposes constraints of jacobi from equation above}.
%\ace{I'm not quite sure why the signs are mentioned in the
%  previous sentence?}

@@

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge.
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of $YZ$-theory \cite{Cheung:2016prv}, J-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-Abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the $J$-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
where the incoming arrow is the abelianized gauge field, and the outgoing arrows the on-shell non-abelian vectors. The propagator is simply:
\be
\propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}
\ee
As one can see, the kinetic term mixes the two gauge fields and has the standard
Feynman rule of $i/k^2$.  We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the abelian polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.

@@

Below we review the four manifestly color-dual theories that we mention in
the text, in particular self-dual Yang-Mills (SDYM) \cite{Monteiro2011pc}, two formulations of NLSM \cite{Cheung:2016prv,Cheung:2020djz,Cheung:2021zvb}, and ZM \cite{Zakharov:1973pp, Cheung:2022mix}.  All
of these theories are closely tied together by the semi-abelian YM
theory in \cref{eq:semiYM}, which gives the cubic sector of pure YM.
The manifestly color-dual equations of motion for SDYM and NLSM are
derived by picking a gauge and then placing a constraint directly on the
field strength. In almost all of these
theories, Bose symmetry is broken or obscured at the Lagrangian level.
Because of the broken Bose symmetry (and the inherited kinetic mixing
of semi-abelian YM), these theories require additional interactions to produce color-dual representations beyond one-loop.

\subsection{Self-Dual Yang-Mills} The first example in the literature
of color-dual Feynman rules is that of Yang-Mills theory in the
self-dual sector (SDYM).  We summarize the appoach of
Ref.~\cite{Monteiro2011pc} for extracting the kinematic algebra.
% \begin{equation}
% \mathcal{L}^{\text{YM}} = -\frac{1}{4}F_{\mu\nu}F^{\mu\nu}
% \end{equation}
To obtain the Lagrangian for self-dual Yang-Mills from standard Yang-Mills, we apply the
self-dual condition that equates the non-abelian $SU(N)$ field
strength,
$F_{\mu\nu} = \partial_\mu A_\nu -\partial_\nu A_\mu + g[A_\mu
,A_\nu]$, to its dual two-form:
\begin{equation}
  \label{SDcon} F_{\mu\nu} = \frac{1}{2}
  \varepsilon_{\mu\nu\rho\sigma}F^{\rho\sigma} \,.
\end{equation}
Solutions to the self-dual condition automatically satisfy the
Yang-Mills equation of motion due to the Bianchi identity:
\begin{equation}
D^\mu F^{\mu\nu} = \frac{1}{2} \varepsilon_{\mu\nu\rho\sigma}D^\mu F^{\rho\sigma} = 0 \,.
\end{equation}
Furthermore, field configurations that satisfy \cref{SDcon} correspond to instantons which are probed by the path integral in the full non-perturbative description of Yang-Mills. To construct a perturbative action, begin by
transforming from Cartesian coordinates to light cone coordinates,
$(t,x,y,z) \rightarrow (u,v,w,\bar{w})$, via:
\begin{equation}
u = t+x,\quad v=t-x, \quad w = y+iz, \quad \bar{w} = y-iz
\end{equation}
which yields the invariant line element:
\begin{equation}
ds^2 = 2(dudv - dwd\bar{w}) = dt^2 - d\vec{x}\cdot d\vec{x} \,.
\end{equation}
Applying this coordinate transformation to the self-dual condition of
\cref{SDcon}, produces three simple constraint equations:
\begin{align} \label{SDLightCone1}
F_{uw} &=0\,,
\\
 \label{SDLightCone2}
 F_{uv} &= F_{w\bar{w}}\,,
\\
 \label{SDLightCone3}
 F_{v\bar{w}} &= 0\,.
\end{align}
Thus, in light-cone gauge, $A_u=0$, we realize the self-dual
conditions in terms of the gauge field as
\begin{align}
  A_u=0 \text{ and } (\ref{SDLightCone1}) \quad&\Rightarrow \quad A_w=0\\
  A_w=0\,,\,A_u=0 \text{ and } (\ref{SDLightCone2}) \quad&\Rightarrow \quad \partial_u A_v = \partial_w A_{\bar{w}} \label{eq:sd-divs}\,.
\end{align}
This last constraint is satisfied by writing the gauge field in
terms of a single degree of freedom which we call $\Psi$:
\begin{equation}
  \label{SDDef}
A_v = \frac{1}{2} \partial_w \Psi, \qquad A_{\bar{w}} = \frac{1}{2} \partial_u \Psi \,.
\end{equation}
Using this definition on the first self-dual constraint equation
yields the following equation of motion for the dynamical scalar
field:
\begin{equation}
  \label{SDEOM}
(\ref{SDLightCone1}) \text{ and } (\ref{SDDef}) \quad  \Rightarrow\quad \Box \Psi + ig [\partial_u \Psi, \partial_w \Psi ]=0 \,.
\end{equation}
Adding back in the anti-holomorphic field $\bar{\Psi}$ as a Lagrange
multiplier, we obtain the following interaction Lagrangian for
self-dual Yang-Mills theory:
\begin{equation}
  \mathcal{L}^{\text{SDYM}} = (\partial \bar{\Psi})(\partial \Psi) -i g \bar{\Psi} [\partial_u \Psi, \partial_w \Psi ] \,.
\end{equation}
The color-ordered ordered cubic Feynman rule for this theory can be
immediately read off as:
\begin{equation}
\cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =X(k_1,k_2) \equiv k_{u,1}k_{w,2} - k_{w,1}k_{u,2} \,.
\end{equation}
In this form, the Lagrangian and Feynman rules are manifestly
color-dual \textit{off-shell}. We can see the effect \cref{SDcon} had
on the Feynman rules of the full Yang-Mills Lagrangian: it simply
decoupled the anti-MHV three-point vertex from the theory, leaving
only the MHV three-point vertex. Before proceeding, it is important to
note that the mass-dimension of the theory appears to differ from that
of Yang-Mills -- we comment on this in detail in \cref{Conventions}.

\subsection{YZ-theory}
Self-dual Yang-Mills involves explicit reliance on the spacetime
dimension, which makes the theory poorly suited for constructing
dimensionally-regulated loop-level amplitudes.  On the other hand,
$YZ$-theory of \cite{Cheung:2016prv} is an honest $D$-dimensional theory that captures the
classical physics of the nonlinear sigma model (NLSM) while
manifesting the duality between color and kinematics.

One can understand $YZ$-theory as a particular dimensional reduction of Yang-Mills in $D=2d+1$ dimensions \cite{Cheung:2017yef}, down to $d$-dimensions. 
Starting with the Yang-Mills Lagrangian of \cref{eq:ym-lag}, one can redefine the gauge fields, $A_M$, in terms of $X$, $Y$ and $Z$ fields:
\begin{align}
X_M &= (X_\mu,0,-iX_\mu) 
\\
Y_M &= (0,Y,0) 
\\
Z_M &= (Z_\mu,0,iZ_\mu) 
\end{align}
where $X$ and $Z$ are $d$-vectors and $Y$ is a scalar field. Considering the conjugate nature of the $XZ$ propagator, in this work we make the replacement $X\rightarrow \bar{Z}$. Plugging this redefinition of the gauge fields in,, we thus obtain the following Lagrangian up to cubic order in the interactions:
\begin{equation}
\mathcal{L}^{\text{YZ}} =\frac{1}{2} (\partial Y)^2 + (\partial Z)(\partial \bar{Z}) - g f^{abc} \left( \bar{Z}_{\mu\nu}Z^{\mu} Z^\nu + [Y,\partial_\mu Y] Z^\mu \right)
\end{equation}
The Feynman rules for this Lagrangian are simply:
\begin{equation}
\Acubic{hgrey0}{}{nhpRed}{}{nhpRed}{fermion2}{} = i (\varepsilon_3 p_2)\qquad \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{} =i(\varepsilon_1 p_{3})(\varepsilon_2\bar{\varepsilon}_3) - i(\varepsilon_2 p_{3})(\varepsilon_1\bar{\varepsilon}_3)
\end{equation}
Note that the pure vector vertex on the left is \textit{exactly} what was obtain from semi-abelian YM in \cref{eq:semiYM}. Using this, we review of the construction of NLSM numerators at tree-level. In this construction, we can define generators of the kinematic algebra as follows
\be\label{eq:FeynmanRuleYYZ}
T^a_{ij}= i \varepsilon_a(p_i-p_j)
\end{equation}
where momentum conservation requires
\begin{equation}
p_a + p_i + p_j =0
\end{equation}
The kinematic half-ladder diagrams then takes on the following concise definition:
\begin{equation}
n^{\text{NLSM}}_{(i|a_1a_2...a_n|j)} = (T^{a_1}T^{a_2}\cdots T^{a_n})_{ij}
\end{equation}
Since there are no pole cancelling factors of $s_{ij} = (p_i+p_j)^2$, this definition of the chiral current algebra is manifestly cubic. Thus, the kinematic structure constants defined in terms of these generators are invariant under generalized gauge freedom. They can be defined implicitly below:
\begin{equation}
[T^a,T^b]_{ij}= F^{a}_{\,b|c}T^c_{ij}
\end{equation}
Given this definition, the Feynman rule associated with kinematic structure constant is simply:
\be\label{eq:FeynmanRuleXZZ}
i F^{a}_{\,b|c} = (\varepsilon_b p_{ab})(\varepsilon_a\bar{\varepsilon}_c) - (\varepsilon_a p_{ab})(\varepsilon_b\bar{\varepsilon}_c) 
\end{equation}
where $p_{ab}=p_a+p_b$ and $\varepsilon$ and $\bar{\varepsilon}$ are the polarizations of the $Z$-vectors particle and its conjugate field, respectively. The gauge-fixed state sum for this theory is simply:
\begin{equation}
\sum_{\text{states}} \varepsilon^{\,\mu}_{(p)}\bar{\varepsilon}^{\,\nu}_{(-p)} = \eta^{\mu\nu}
\end{equation}
Notice that the vector state sum is gauge fixed since the $YZ$ model explicitly chooses Lorenz gauge for the $Z$ particles, $\partial_\mu Z^\mu=0$. To recover NLSM amplitudes from this these kinematic structure constants at tree-level, we simply plug in the following on-shell polarizations for the $Z$ and $\bar{Z}$ particles:
\be\label{eq:onShellZStates}
\varepsilon^\mu_{(p)} = p^\mu \qquad \bar{\varepsilon}^\mu_{(p)} = \frac{q^\mu}{pq}
\end{equation}
where $q^2=0$ is some null reference momentum. With this, we can define tree-level pion scattering in two equivalent ways:
\begin{equation}
A^{\text{NLSM}} = A(...,Y,...,Y,...) = A(...,\bar{Z},...) 
\end{equation}
where the ellipses denote additional on-shell $Z$-particles. Subject to a particular gauge choice, the kinematic numerators in latter definition for pion scattering is equivalent to those $J$-theory, first written down by Cheung and one of the authors \cite{Cheung:2021zvb}. In this paradigm, the on-shell $\bar{Z}$ state is equivalent to the root leg appearing in the color-dual $J$-theory numerators.
